---
params:
  projectDir: 'D:\LP_Files\RStudio_Files\trail_counter_2011_2020\scripts\IMD_DRR_Template'       
title: "Coastal Area Visitation Trends 2011-2020"
author:
  - name: "Lauren L.M. Pandori & Linh Anh Cat"                                      
editor_options:
  chunk_output_type: inline
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
link-citations: yes
output:
  bookdown::html_document2:
    df_print: kable
    fig_caption: true
    dev: svg
    highlight: haddock
    keep_md: no
    smart: no
    theme: journal
    css: "common/journalnps.min.css"
    toc: no
    number_sections: false
    includes:
        before_body:
          - "common/header.html"
  bookdown::word_document2:
    df_print: kable
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: haddock
    number_sections: false
---
```{r md setup, include=FALSE}
library(flextable)

# set knitr options for all chunks
knitr::opts_chunk$set(
   root.dir = params$projectDir,  # from YAML parameter, knitr instead of setwd()
   echo = TRUE,
   comment = " ",
   dev = "svg",
   fig.path = "figures/",
   tidy.opts = list(width.cutoff = 60),
   tidy = TRUE
   )

# if ggplot, update theme to default to centered titles
if ("ggplot2" %in% .packages()) {
   theme_update(plot.title = element_text(hjust = 0.5))
}

# Set up table template
NPS_theme <- function(x, ...) {
    x <- colformat_double(x, big.mark = '', decimal.mark = ".", digits = 0)
    x <- colformat_int(x, )
    x <- colformat_date(x,fmt_date = "%Y-%m-%d")
    x <- set_table_properties(x, layout = "fixed")
    x <- border_remove(x)
    std_border <- fp_border_default(width = 1, color = "black")
    x <-hline_bottom(x,part="body")
    x <-hline_bottom(x,part="header")
    x <-hline_top(x,part="header")
    x <-bold(x,bold=TRUE,part="header")
    x <-set_table_properties(x, width = 0, layout = "autofit")
    x <-align_nottext_col(x, align = "right", header = TRUE, footer = TRUE)
    x <-align_text_col(x, align = "left", header = TRUE, footer = TRUE)
    x <-valign(x,valign="bottom",part="header")
    x <-valign(x,valign="top",part="body")
}

set_flextable_defaults(
  font.family = "Arial", 
  font.size = 9,
  font.color = "black",
  theme_fun = NPS_theme,
  )

# set working directory (from YAML)
setwd(params$projectDir)

# figure text wrapping 
# https://stackoverflow.com/questions/54100806/r-markdown-how-do-i-make-text-float-around-figures

defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

```{r data setup, include=FALSE}

# run tm_data_analysis script to set up environment
devtools::source_url('https://raw.githubusercontent.com/llmpandori2/trail_counter_2011_2020/main/scripts/tm_data_analysis.R')

```
<hr>
### Summary
Cabrillo National Monument (San Diego, CA, USA) is a pint-sized unit of the National Park System (160 terrestrial acres) that attracted an average of 850,720 annual visitors 2011-2020. The coastal area of the park contains well-preserved and sensitive natural resources, including sea cliffs, coastal and maritime sage scrub, and rocky intertidal habitat. Understanding visitor use patterns in this area can guide management decisions such as congestion planning and allocation of recreation fee project money toward staffing the area. Infrared trail counters (TrailMaster™ 1550, USA) were deployed adjacent to two entrances to the Coastal Trail at the Tidepool Parking and Coast View Parking areas to capture visitation patterns from 2011-2020.

#### Low tide days don't result in increased visitation
Despite the tidepools being a popular and desired visitor experience, tide level was not correlated with coastal area visitation. Despite these findings, increased staff and volunteer presence is needed during low tides to steward sensitive habitat. 

#### Coastal area visitation has increased over time
``````{r figure1, echo = F, warning = F, message = F, fig.height = 3, fig.width = 5, out.width = '50%', out.extra = 'style = "float:right; padding 5px"'}

visit_est %>%
  pivot_wider(names_from = 'Location', values_from = 'Visitors') %>%
  mutate(`Coastal area` = round(`Lot 1` + `Lot 2`, digits = 0)) %>%
  mutate(`% of Visitors Exploring Coastal Area` = (`Coastal area`/`Entrance station`)) %>%
  rename(`Coastal Area` = `Coastal area`, `Entrance Station` = `Entrance station`) %>%
  ggplot(data = .) +  
    geom_col(mapping = aes(x = Year, y = (`% of Visitors Exploring Coastal Area`*1000000)), fill = 'gray80') +
    geom_line(mapping = aes(x = Year, y = `Coastal Area`), 
              color = cal_palette('tidepool')[1], size = 2) +
    geom_text(mapping = aes(x = 2011, y = `Coastal Area`[1] + 150000, label = 'Coastal area visitors'), 
              color = cal_palette('tidepool')[1], hjust = 0) + 
    geom_text(mapping = aes(x = 2011, y = `Entrance Station`[1] + 210000, label = 'Entrance station visitors'), 
              color = cal_palette('tidepool')[5], hjust = 0) +
    geom_text(mapping = aes(x = 2011, y = 150000, 
                            label = 'Percent of all visitors exploring coastal area'),
              color = 'gray30', hjust = 0) + 
    geom_line(mapping = aes(x = Year, y = `Entrance Station`), 
              color = cal_palette('tidepool')[5], size = 2) + 
 
    scale_y_continuous(labels = scales::comma, 
      sec.axis=sec_axis(~.*0.0001, name = 'Coastal area visitors (%)')) + 
    scale_x_continuous(breaks = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)) +
    ylab('Annual visitors') +
    #scale_x_date(breaks = '1 year') + 
    lltheme_light +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.95),
          text = element_text(color = 'black', size = 11))
  
```
While total visitation to the park has fluctuated over time, visitation to the coastal area has increased steadily during the study period (2011-2020). Overall visitation increased 2011-2016 up to the National Park Service Centennial, but decreased after 2017 in conjunction with increasing entrance fees and the COVID-19 Pandemic. Despite changes in entrance station visitation, the number of visitors to the coastal area increased 2011-2020. **In 2020, 73% of park visitors explored the coastal area, compared to 37% in 2011.**

#### Weekends attract more visitors
Weekend days (Saturday and Sunday) had the highest visitation, with an average of 1246 visitors per day entering the Coastal Trail from the Tidepool Parking lot and 370 visitors per day from the Coast View Parking lot. Weekend-edge days (Monday and Friday) had significantly lower visitation than weekend days, but higher visitation than midweek days (Tuesday through Thursday).

Across all days of the week, visitor activity is high at park opening (9:00 AM), drops mid-day, and peaks at closing time (4:30-5:00 PM).

#### Visitation increases most holidays
```{r echo = F, warning = F, message = F, fig.height = 5, fig.width = 7, out.width = '50%', out.extra = 'style = "float:right; padding 5px"'}

ggplot(data = holiday_test) + 
  geom_col(mapping = aes(y = mean_dif, x = fct_rev(holiday), fill = pval_sig)) + 
  geom_errorbar(mapping = aes(x = fct_rev(holiday), ymin = (mean_dif - se_dif), 
                              ymax = (mean_dif + se_dif), width = 0.3),
                color =  'gray48') + 
  scale_fill_manual(values = c('gray85', cal_palette('tidepool')[1])) + 
  geom_text(mapping = aes(x = 14, y = -100, label = 'Fewer visitors'), 
            color = 'black', hjust = 0.95, size = 3.5) + 
  geom_text(mapping = aes(x = 14, y = 100, label = 'More visitors'), 
            color = 'black', hjust = 0, size = 3.5) + 
  geom_text(data = filter(holiday_test, pval_sig == TRUE),
            mapping = aes(x = fct_rev(holiday), y = (mean_dif + se_dif + 100), 
                          label = '*'), color = 'black') +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'black') + 
  coord_flip(xlim = c(1,14)) +
  scale_y_continuous(breaks = seq(-1000,1000, by = 500), limits = c(-1200,1200)) +
  ylab('Difference in visitation') + 
  xlab('Holiday') +
  facet_wrap(~lot, labeller = as_labeller(c('Lot 1' = 'Tidepool Parking',
                                            'Lot 2' = 'Coast View Parking'))) + 
  lltheme_light + 
  theme(legend.position = 'none',
        panel.grid = element_blank(),
        text = element_text(color = 'black', size = 11),
        axis.text = element_text(color = 'black'),
        axis.text.x = element_text(angle = 45, hjust = 1))
```
Apart from New Year’s Day, Christmas Day and Veterans Day, all holidays and holiday-adjacent days considered had significantly higher visitation than fourteen days after (as indicated by stars on figure). While there is some overlap between fee-free days and holidays, **there was no strong pattern between fee-free days and visitation.**

On average, holidays and holiday-adjacent days attract over 800 additional visitors to the coastal area! 

### Recommendations
Findings indicate that peak staffing of Volunteers-in-Parks (VIPs) and park staff should occur on weekends and holidays, regardless of tide level. We recommend to double staffing on weekends and holidays, with an even higher multiplier on holiday weekends.


